Paradigms in Physics 2.0

### 2018-09-29 proposal.tex
Changes by Emily

\newcommand\corinnesays[1]{\textcolor{red}{[\sc CAM: {\em#1}]}}
\newcommand\davidsays[1]{\textcolor{blue}{[\sc DR: {\em#1}]}}

@@

In 2016, the Department of Physics at Oregon State University began a
process to revise our Paradigms in Physics curriculum for physics
majors.  We began with a colloquium to inform the department of our
plans and request their assistance, followed by a
survey of students and faculty as well as individual interviews with
the faculty teaching each course.  As we developed a plan to
address student- and faculty-identified challenges in the curriculum,
we met with each faculty member individually to explain and refine our
proposal, which was unanimously approved by the faculty.  Major
changes include major changes to several courses (math
methods, computational physics, modern physics, electronics, and
classical mechanics), including the introduction of two sophomore-year
courses designed specifically to help prepare students for their
upper-division courses.

This paper will describe both the process of the reform and its
outcome, but the emphasis of the paper and the research aspect of it
will be on the process by which we developed the changes and arrived
at consensus.  This is paper is thus on the coarse-grained picture of
how one department acheived a broad-scale reform of its curriculum,
which we expect will serve as a valuable balance to the more
fine-grained view that is needed for classroom activities.  In
addition to the Paradigms 2.0 reform, we will briefly introduce the
original effort that began in 1996, and give a bit more detail on how
that reform was maintained and refined during the intervening decades.

This story of how we achieved a dramatic curriculum change will be
useful for other departments considering a reform of their curriculum.
Few departments embark on such an endeavor, and publishing a blueprint
for how we managed this will be helpful for others seeking to reform
their curriculum.

\paragraph*{Theoretical basis for the work}
The framework for this {\bf empirical study} was adapted from studies
of organizational decision-making\cite{beach2006leadership}.  Like the
process that led to Paradigms 1.0, the Paradigms 2.0 process is a
``shared vision'' type of change\cite{henderson2010}. Changes to the
course structures of the physics curriculum emerged from discussions
among the members of the department. The Paradigms~2.0 process began
in the Winter of 2016, after the faculty agreed to the process during
the previous year.

\paragraph*{Research questions}
\begin{itemize}
\item What was the culture of the department when the faculty initiated revising the upper-level physics curriculum? In
  particular, what were the participants' beliefs about teaching and
  learning?
\item What motivated making changes and how did the faculty form a
  coherent vision for the new program?
\item How did the faculty develop and communicate plans for this new
  curriculum?
\item How have they changed instruction based upon the proposals that
  the faculty approved with an unanimous vote?
\end{itemize}
\paragraph*{Research methods}
We performed interviews of faculty before, during and after the
Paradigms 2.0 process, gave surveys to students and faculty, and had
an observer take notes during all Paradigms 2.0 meetings.

>>>>

\newcommand\evzsays[1]{\textcolor{red}{[\sc EvZ: {\em#1}]}}
\newcommand\davidsays[1]{\textcolor{blue}{[\sc DR: {\em#1}]}}

@@

In 2016, the Department of Physics at Oregon State University began a
process to revise our Paradigms in Physics curriculum for physics
majors. Paradigms~2.0 began in the Winter of 2016, after the faculty agreed to undertake this reform during the previous year. The process included frequent meetings by a small committee of faculty volunteers, a colloquium to communicate plans and to request assistance, an online survey of students and faculty to identify issues, individual interviews with the faculty teaching each course, and discussions with focus groups of students and instructional faculty.

As we developed ideas to address challenges in the curriculum, we met with each faculty member individually to explain and refine our proposal. Major changes included revising a series of courses, eliminating others, re-conceptualizing how to integrate just-in-time mathematics instruction, and introducing two sophomore-year courses designed specifically to help prepare students for their upper-division courses. The proposal was unanimously approved by the faculty and then implemented during 2016-2017.

This paper will describe both the process of the reform and its
outcome. It will provide a coarse-grained picture of how one department achieved a broad-scale reform of its upper-division curriculum, which can serve as a valuable balance to the more fine-grained view that is needed for refining classroom activities or revising a single course.  In addition to the Paradigms 2.0 reform, we will briefly introduce the original effort that began in 1996 \cite{manogue2001paradigms,manogue2003paradigms}, and summarize how that reform was maintained and refined during the intervening decades. Few departments embark on such a long-term endeavor and publishing a report on how we managed this may be helpful for others seeking to reform their curriculum.

\paragraph*{Theoretical basis for the work}
The framework for this {\bf empirical study} was adapted from studies
of organizational decision-making\cite{beach2006leadership}.  Like the
process that led to Paradigms 1.0, the Paradigms 2.0 process is a
``shared vision'' type of change\cite{henderson2010}. Changes to the
course structures of the physics curriculum emerged from on-going collaborative discussions among the members of the department.

\paragraph*{Research questions}
\begin{itemize}
\item What was the culture of the department when the faculty initiated revising the upper-level physics curriculum? In
  particular, what were the participants' beliefs about teaching and
  learning?
\item What motivated making changes and how did the faculty form a
  coherent vision for the revised program?
\item How did the faculty develop and communicate plans for this revised
  curriculum?
\item How have they changed instruction based upon the proposals that
  the faculty approved with an unanimous vote?
\end{itemize}
\paragraph*{Research methods}
Data included audio recordings of interviews with faculty before, during and after the Paradigms 2.0 process; responses to an online survey of students and faculty; audio recordings, notes, and some photos during Paradigms 2.0 Committee meetings, Upper Division Curriculum Committee meetings, and some faculty meetings; and copies of relevant documents. Analysis included a quantitative summary of survey data and qualitative interpretations of what happened while initiating and sustaining the Paradigms in Physics program (1996-2015), re-envisioning the Paradigms Program during 2016, and implementing Paradigms 2.0 during 2016-2017 and beyond.
